Miniproject - Combinatorics

### 2021-12-02 Combinatorics/submission.tex
Finished combinatorics part 4

\section{Part IV}
\begin{example}
    Let $\Pi = \{ \pi \subseteq \powerset(n) \ | \ \text{$\pi$ partition} \}$ be the poset of partitions ordered by refinement.
    Then $\Pi$ is a graded poset with rank function
    \begin{equation*}
        r: \Pi \to \{ 0, ..., n - 1 \}, \quad \pi \mapsto n - |\pi|
    \end{equation*}
\end{example}
\begin{proof}
    First of all, have $1 \leq |X| \leq n$ for all subsets $X \subseteq \powerset(n)$, hence the function $r$ is well-defined.
    Next we show that for all $x < y$ have $r(x) < r(y)$.
    However, if $x$ is a proper refinement of $y$, then clearly $|x| > |y|$, so $r(x) < r(y)$.

@@

    Note that this is the partition into \emph{exactly} $k$ sets; 
    I am not completely sure what the exercise description tells us to show, but if it is about the number of ways of decomposing $[m]$ into \emph{at most} $k$ sets, then have
    \begin{equation*}
        \Sigma S(m, k)(2) = 2^{m - 1} \quad \text{and} \quad \Sigma S(m, k)(3) = 3^{m - 1} = \frac 1 2 3^{m - 1} - \frac 1 2
    \end{equation*}
    for $m \geq 3$.
    Here $\Sigma S(m, k) = \sum_{l \leq k} S(m, l)$ is the number of partitions of $[m]$ into at most $k$ sets.
\end{example}
\begin{proof}
    By Prop.~\ref{prop:partitions_recursion} have that for $m \geq 1$ we have $S(m, 1) = 1$.
    Further, for $m \geq 2$ have
    \begin{equation*}
        S(m, 2) = \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 = 2^{m - 1} - 1
    \end{equation*}
    Applying this once more yields for $m \geq 3$ that
    \begin{align*}
        S(m, 3) =& \sum_{n = 1}^{m - 1} {m - 1 \choose n} (2^{n - 1} - 1) = \frac 1 2 \sum_{n = 1}^{m - 1} {m - 1 \choose n} 2^n - \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 \\
        =&\frac 1 2 \sum_{n = 0}^{m - 1} {m - 1 \choose n} 2^n - \frac 1 2 - \sum_{n = 0}^{m - 1} {m - 1 \choose n} 1 + 1 \\
        =& \frac 1 2 (1 + 2)^{m - 1} - 2^{m - 1} + \frac 1 2 = \frac 1 2 (3^{m - 1} - 2^m + 1)
    \end{align*}
\end{proof}

>>>>

\section{Part IV}
\begin{example}
    Let $\Pi_m = \{ \pi \subseteq \powerset(n) \ | \ \text{$\pi$ partition} \}$ be the poset of partitions ordered by refinement.
    Then $\Pi$ is a graded poset with rank function
    \begin{equation*}
        r: \Pi \to \{ 0, ..., n - 1 \}, \quad \pi \mapsto n - |\pi|
    \end{equation*}
\end{example}
\begin{proof}
    First of all, have $1 \leq |X| \leq n$ for all subsets $X \subseteq \powerset(n)$, hence the function $r$ is well-defined.
    Next we show that for all $x < y$ have $r(x) < r(y)$.
    However, if $x$ is a proper refinement of $y$, then clearly $|x| > |y|$, so $r(x) < r(y)$.

@@

    Note that this is the partition into \emph{exactly} $k$ sets; 
    I am not completely sure what the exercise description tells us to show, but if it is about the number of ways of decomposing $[m]$ into \emph{at most} $k$ sets, then have
    \begin{equation*}
        \Sigma S(m, k)(2) = 2^{m - 1} \quad \text{and} \quad \Sigma S(m, k)(3) = 3^{m - 1} = \frac 1 2 3^{m - 1} + \frac 1 2
    \end{equation*}
    for $m \geq 3$.
    Here $\Sigma S(m, k) = \sum_{l \leq k} S(m, l)$ is the number of partitions of $[m]$ into at most $k$ sets.
    \begin{proof}
        By Prop.~\ref{prop:partitions_recursion} have that for $m \geq 1$ we have $S(m, 1) = 1$.
        Further, for $m \geq 2$ have
        \begin{equation*}
            S(m, 2) = \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 = 2^{m - 1} - 1
        \end{equation*}
        Applying this once more yields for $m \geq 3$ that
        \begin{align*}
            S(m, 3) =& \sum_{n = 1}^{m - 1} {m - 1 \choose n} (2^{n - 1} - 1) = \frac 1 2 \sum_{n = 1}^{m - 1} {m - 1 \choose n} 2^n - \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 \\
            =&\frac 1 2 \sum_{n = 0}^{m - 1} {m - 1 \choose n} 2^n - \frac 1 2 - \sum_{n = 0}^{m - 1} {m - 1 \choose n} 1 + 1 \\
            =& \frac 1 2 (1 + 2)^{m - 1} - 2^{m - 1} + \frac 1 2 = \frac 1 2 (3^{m - 1} - 2^m + 1)
        \end{align*}
    \end{proof}
    Let $m$ be even and consider $\mathcal{A} \subseteq \Pi_m$ of partitions of $[m]$ into two equally sized sets.
    Then
    \begin{equation*}
        |\mathcal{A}| = \frac 1 2 {m \choose m/2}
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{equation*}
            f: [m]^{(m/2)} \to \mathcal{A}, \quad A \mapsto \{ A, [m] \setminus A \}
        \end{equation*}
        Then for $A$ and $B \neq A, [m] \setminus A$ have that $\{ A, [m] \setminus A \} \neq \{ B, [m] \setminus B \}$.
        Conversely, $A$ and $[m] \setminus A$ have the same image under $f$.
        This shows that $f$ is 2-to-1 and thus 
        \begin{equation*}
            |\mathcal{A}| = \frac 1 2 \left| [m]^{(m/2)} \right| = \frac 1 2 {m \choose m/2}
        \end{equation*}
    \end{proof}
    Further have
    \begin{equation*}
        |\partial \mathcal{A}| = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{align*}
            g: \underbrace{\{ (A, B, C) \ | \ A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m \}}_{ =: \mathcal{G} } \to \partial \mathcal{A}, \quad (A, B, C) &\mapsto \{ A, B, C \}
        \end{align*}
        Note that $g$ is well-defined, as for $A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m$ have that $\pi := \{ A, B, C \}$ is a partition in $L_{m - 3}$ that refines the partition $\{ A, B \cup C \} \in \mathcal{A} \subseteq L_{m - 2}$.
        Further, it is easy to see that $g$ is surjective, as every partition $\pi \in \partial \mathcal{A}$ satisfies $|\pi| = 3$ and has some $A \in \pi$ with $|A| = m/2$.

        To complete the proof, we investigate to what ``extend $g$ is injective''.

        Assume $g(A, B, C) = g(A', B', C')$. As $B, C, B', C' \neq \emptyset$ and $|B \cup C| = |B' \cup C'| = m/2$, we see that $A = A'$.
        Further, we must then have that $\{ B, C \} = \{ B', C' \}$, hence $(A, B, C) = (A', B', C')$ or $(A, B, C) = (A', C', B')$ (clearly $B \neq C, B' \neq C'$).
        This shows that the map $g$ is 2-to-1.

        Using this, we find
        \begin{align*}
            |\partial \mathcal{A}| =& \frac 1 2 |\mathcal{G}| = \frac 1 2 \sum_{A \in [m]^{(m/2)}} |\{ (B, C) \ | \ \{ B, C \} \ \text{partition of} \ [m] \setminus A \}| \\
            =& \frac 1 2 \sum_{A \in [m]^{(m/2)}} 2 S(m/2, 2) = \sum_{A \in [m]^{(m/2)}} 2^{m/2 - 1} - 1 = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
        \end{align*}
    \end{proof}
    We can now plug this into the second condition of Prop.~\ref{prop:characterization_lym} to see
    \begin{align*}
        \frac {|\partial \mathcal{A}|} {|L_{n - 3}|} &= \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {S(m, 3)} = \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {\frac 1 2 \left( 3^{m - 1} - 2^m + 1 \right)} \\
        &\quad \sim \frac {2^{m + m/2}} {\sqrt{n}(3^{m - 1} - 2^m)} \sim \frac {3 \left( 2^{3/2} \right)^m} {\sqrt{n} 3^m} = \frac 3 {\sqrt{n}} c^m
    \end{align*}
    and
    \begin{align*}
        \frac {|\mathcal{A}|} {|L_{n - 2}|} &= \frac {\frac 1 2 {m \choose m/2}} {S(m, 2)} = \frac {{m \choose m/2}} {2^m - 2} \sim \frac {2^m} {\sqrt{n}2^m} = \frac 1 {\sqrt{n}}
    \end{align*}
    for some $0 < c < 1$ (here $\sim$ means asymptotically equivalent as $m \to \infty$).
    In particular, find for sufficiently large $m$ that
    \begin{equation*}
        \frac {|\partial \mathcal{A}|} {|L_{n - 3}|} < \frac {|\mathcal{A}|} {|L_{n - 2}|}
    \end{equation*}
    and so the conditions of Prop.~\ref{prop:characterization_lym} are not satisfied.
\end{example}
